Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-29 paper/paper.tex
heuristic 5 histogram and comment

\begin{figure*}[t!]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-contents}Number of content options.}
\end{subfigure}% <---- don't forget this %
%\vspace{1ex}
%\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-rule-size}Rule sizes.}
\end{subfigure}%
%\vspace{1ex}
\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-3ex}  
  \caption{\label{fig:distribution-strings}Strings used as parameters of content options.}
\end{subfigure}%
%\vspace{1ex}
%\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.}
\end{subfigure}%
\vspace{-1ex}
  \caption{Distribution of features.}
\end{figure*}

@@

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).

>>>>

\begin{figure*}[t!]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-contents}Number of content options.}
\end{subfigure}% <---- don't forget this %
%\vspace{1ex}
%\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,11) (1,1301) (2,4910) (3,3662) (4,3250) (5,5353) (6,2911) (7, 1337) (8, 1246) (9,763) (10, 495)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-rule-size}Rule sizes.}
\end{subfigure}%
%\vspace{1ex}
\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={GET, POST, SELECT, UNION, id=, DELETE, UPDATE, INSERT, 200, /index.php?},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(GET,2007) (POST,1570) (SELECT,747) (UNION,461) (id=,408) (DELETE,390) (UPDATE,387) (INSERT, 386) (200, 352) (/index.php?, 273)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-3ex}  
  \caption{\label{fig:distribution-strings}Strings used as parameters of content options.}
\end{subfigure}%
%\vspace{1ex}
%\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-1ex}
  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.}
\end{subfigure}%
%\vspace{1ex}
\\
\begin{subfigure}{.5\textwidth}
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={content, flow, pcre, flowbits, threshold, dsize, itype, icode, flags, tls.fingerprint},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(content,25699) (flow,23159) (pcre,10181) (flowbits,3049) (threshold,2300) (dsize,882) (itype,155) (icode, 108) (flags, 97) (tls.fingerprint, 61)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-3ex}  
  \caption{\label{fig:distribution-strings}Option Frequency.}
\end{subfigure}%
\vspace{-1ex}
  \caption{Distribution of features.}
\end{figure*}

@@

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).
\Gui{A quinta heuristica que utilizamos considera a popularidade das opcoes. Primeiro pegamos a quantidade de vezes que uma opcao eh utilizada (sem considerar repeticoes numa unica regra). Dividimos o valor encontrado para cada opcao pelo maior valor encontrado, encontrando sempre valores de no maximo 1. Entao na regra nos somamos os valores de cada opcao, dividido pelo numero de opcoes, chegando a um numero de no maximo 1, que representa a popularidade das opcoes na regra.
\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}
